Aligning marine species range data to better serve science and conservation

### 2016-02-05 ms_3_methods.Rmd

* Possible analyses:
    * run AquaMaps (new data) with threshold at 0%
        * already done in original paper for old data, but redo it to compare to AM + IUCN scenario below
    * run with AquaMaps only, but filtering out data-poor species?
    * run with IUCN maps only in place of AquaMaps?
    * run with AquaMaps and IUCN both, using 0% threshold; which preferred? 
        * Should we clip corals (and coral-associated taxa) to 200 meter bathymetry, just for fun?
* Which aspects of the analysis should we focus on?:
    * overall vs broken down by taxonomic groups? (see paper, figure 1)
    * spatial distribution of gap species (see paper, figure 2)

>>>>

* Possible analyses: <!--BSH: priorities are 1 (AM only), 3 (IUCN only) and 4 (AM + IUCN).-->
    * run AquaMaps (new data) with threshold at 0%
        * already done in original paper for old data, but redo it to compare to AM + IUCN scenario below
    * run with IUCN maps only in place of AquaMaps? _<!--BSH: this seems a bit silly, since AM has so many more species, but worth doing to make the point, I think.-->_
    * run with AquaMaps and IUCN both, using 0% threshold; which preferred? _<!--BSH: I think this would be good to do, with AM as priority.  We can then compare to AM only, as you note above.  This tells us exactly what is gained by including the additional 1500 species (or whatever) that are only in IUCN.-->_
* Which aspects of the analysis should we focus on?: _<!--BSH: I’m guessing it will be hard to see differences in the maps (except for the IUCN only example), so probably best to focus on figure 1 presentation of results.-->_
    * overall vs broken down by taxonomic groups? (see paper, figure 1)
    * spatial distribution of gap species (see paper, figure 2) <!--

### 2016-02-05 ms_4_figs.Rmd
updated figures; added word count; etc

@@

### Fig. 1

<a id = 'fig1'></a>

@@

### Figure 2

@@

> __Fig. 2.__ (A) Distribution alignment (overlap of smaller range within larger) versus extent ratio (the ratio of smaller range area to the larger range area) for `r n_spp_both` species included in both IUCN and AquaMaps datasets.  The upper right quadrant (quadrant 1) comprises species whose maps largely agree (better than median value) in both spatial distribution and the extent of described ranges
(n = `r q1_count`; `r round(q1_count/nrow(spp_map_pairs) * 100, 1)` %).
> The upper left quadrant (quadrant 2) comprises species whose maps agree well in distribution, but disagree in extent
(n = `r q2_count`; `r round(q2_count/nrow(spp_map_pairs) * 100, 1)` %).
> The lower right quadrant (quadrant 3) includes species for which the paired maps generally agree in range extent, but disagree on where those ranges occur
(n = `r q3_count`; `r round(q3_count/nrow(spp_map_pairs) * 100, 1)` %).
> The lower left quadrant (quadrant 4) indicates species for which the map pairs agree poorly in both area and distribution
(n = `r q4_count`; `r round(q4_count/nrow(spp_map_pairs) * 100, 1)` %).  
(B) Alignment breakdown of paired-map species by taxonomic group. _is it distracting to use order names for some groups and not others?  Some of the IUCN taxonomic groups had pretty long names ("acanthuridae" is way shorter than "surgeonfish, tangs, and unicornfish")_

@@

___NOTE: These figures still need to be recreated using the latest and greatest data!___ _currently they use data generated in November or so, in ohiprep._

@@

> * Scenario 1 shows the effect of reducing the presence threshold for AquaMaps presence.  Reducing the threshold increases the apparent range of a species; the slight decrease in average score suggests increased spatial representation of threatened species.
> * Scenario 2 shows the effect of prioritizing AquaMaps data over IUCN, while maintaining the same presence threshold.  This will have different effects depending on the species; in general, AquaMaps ranges are smaller than IUCN ranges, so most overlapping species will see a decrease in represented range.  The slight bump in mean score may indicate a small increase in spatial representation of low-risk species, a small decrease in spatial representation of high-risk species, or more likely a combination of both.
> * Scenario 3 shows the effect of prioritizing AquaMaps data over IUCN, while eliminating the presence threshold.  Just as a presence threshold of zero in scenario 1 drives a decrease in average score relative to the baseline, the zero threshold in scenario 3 drives a decrease in scores relative to scenario 2.  The large decrease indicates that within the set of paired-map species, a zero threshold greatly increases the spatial representation of high-risk species relative to low-risk species.

>>>>

Figures: Calculated at 180 characters per cm in height for one column and 360 characters per cm in height for two columns; a figure wider than 8.7 cm is two columns wide.

@@

### Fig. 1 <!--can you put N = XXX above each bar? Would make it a little easier to read compared to within the bar.-->
<a id = 'fig1'></a>

@@

### Figure 2 <!--Ditto of Fig. 2B – can you move the N = XX values to be outside the bars (to the right) instead of within them?-->

@@

_CCO: a lot of this is duplicated in body… tell the story more briefly here; BSH: Yes, make very brief, probably cutting most of this._
> __Fig. 2.__ (A) Distribution alignment (overlap of smaller range within larger) versus extent ratio (the ratio of smaller range area to the larger range area) for `r n_spp_both` species included in both IUCN and AquaMaps datasets.  The upper right quadrant (quadrant 1) comprises species whose maps largely agree (better than median value) in both spatial distribution and the extent of described ranges
(n = `r q1_count`; `r round(q1_count/nrow(spp_map_pairs) * 100, 1)` %).
> The upper left quadrant (quadrant 2) comprises species whose maps agree well in distribution, but disagree in extent
(n = `r q2_count`; `r round(q2_count/nrow(spp_map_pairs) * 100, 1)` %).
> The lower right quadrant (quadrant 3) includes species for which the paired maps generally agree in range extent, but disagree on where those ranges occur
(n = `r q3_count`; `r round(q3_count/nrow(spp_map_pairs) * 100, 1)` %).
> The lower left quadrant (quadrant 4) indicates species for which the map pairs agree poorly in both area and distribution
(n = `r q4_count`; `r round(q4_count/nrow(spp_map_pairs) * 100, 1)` %).  
(B) Alignment breakdown of paired-map species by taxonomic group. _is it distracting to use order names for some groups and not others?  Some of the IUCN taxonomic groups had pretty long names ("acanthuridae" is way shorter than "surgeonfish, tangs, and unicornfish")_ _BSH: No, but just be sure to note this difference in the legend._ ___NOTE: including the order names might imply the whole order is covered... not necessarily so?  Use IUCN taxo groups!___

@@

___NOTE: These figures still need to be recreated using the latest and greatest data!___ _currently they use data generated in November or so, in ohiprep._ | _BSH: It will be worth noting in the main text of the paper the range of values, i.e., that individual countries can see differences from -12 to +6 (depending on scenario).  Those are pretty big differences within a country, even if the overall global number doesn’t change more than a few points._

@@

> * Scenario 1 shows the effect of reducing the probability threshold for AquaMaps presence.  Reducing the threshold increases the apparent range of a species; the slight decrease in average score suggests increased spatial representation of threatened species. _CCO: Reference how this more accurately tracks the definition of IUCN “limits of distribution”?  BSH: I don’t think you need to directly reference this comparison, but these scenarios are helpful to understand what’s going on, so they’ll need some explanation/implications._
> * Scenario 2 shows the effect of prioritizing AquaMaps data over IUCN, while maintaining the same presence threshold.  This will have different effects depending on the species; in general, AquaMaps ranges are smaller than IUCN ranges, so most overlapping species will see a decrease in represented range.  The slight bump in mean score may indicate a small increase in spatial representation of low-risk species, a small decrease in spatial representation of high-risk species, or more likely a combination of both. _CCO: Too much detail here? Just get the main point – probably a mixture of more/less good/bad – can’t draw much of a conclusion other than that scores change all over.  BSH: You’ll need a sentence or two explaining this result, but keep it brief.  Extra text, if deemed useful, can be put into the SOM._
> * Scenario 3 shows the effect of prioritizing AquaMaps data over IUCN, while eliminating the presence threshold.  Just as a presence threshold of zero in scenario 1 drives a decrease in average score relative to the baseline, the zero threshold in scenario 3 drives a decrease in scores relative to scenario 2.  The large decrease indicates that within the set of paired-map species, a zero threshold greatly increases the spatial representation of high-risk species relative to low-risk species. _similar note as scenario 2_
